---
title: "WR depth per scaffold"
author: "Fátima S. Barreiro"
date: "April 19, 2020"
output: html_document
---


Load required libraries
```{r}
library(ggplot2)
library(scales)
library(reshape2)
library(sweep)
library(cowplot)
library(ggpubr)
```

### Choice of scaffolds from the reference assembly based on size

Total length of the assembly
```{r}
# read in scaffold info
fai = read.table("cerSim1.fasta.fai")
scaffolds = as.data.frame(fai[,1:2])

# sort scaffolds by size
ordered_scaffolds = scaffolds[order(scaffolds$V2),]

# calculate cumulative sum of scaffold sizes 
ordered_scaffolds[, 3] <- cumsum(as.numeric(ordered_scaffolds[, 2]))

# total length of assembly
tail(ordered_scaffolds$V3, 1)
```



Calculating the proportion of the genome where we find informative sites
- 63 autosomal scaffolds minus the masked regions
```{r}
1185902722/2464367180
```


Distribution of scaffold sizes
```{r}
Y = ggplot(ordered_scaffolds, aes(V2, V3))
Y = Y + geom_point(colour="royalblue4", size=2, alpha=.8) + theme_bw()
Y = Y + geom_vline(xintercept=1.3e7, colour="darkorange") 
Y = Y + scale_x_continuous(labels = function(x) format(x/1000000, digits=3))
Y = Y + scale_y_continuous(labels = function(x) format(x/1000000, digits=3))
Y = Y + xlab("Scaffold size (Mbp)") + ylab("Cumulative sum of scaffold length (Mbp)")
#Y = Y + theme(axis.text=element_text(size=6, colour="black"), axis.title=element_text(size=7))

Y
```




######################################################################

### Test of depth to identify scaffolds from sex chromosomes among the n = 66 > 13 Mbp

Read in files
```{r}
# retrieve files
Depthfiles66 <- read.table('WR_56I_13Mb.Depth', header=F)
```

```{r}
# initialize empty dataframe to collect the data and the new columns
B = NULL
```

Expand the data frame with the pertinent calculations
```{r}
for (i in levels(Depthfiles66$V5)) {
  
  # read in files
  b <- Depthfiles66[Depthfiles66$V5 == i, ]
  colnames(b) = c("Scaffold", "InitialPosition", "EndPosition", "DepthCounts", "Sample")
  
  # compute total depth of coverage per individual
  TotalDepthInd = sum(as.numeric(b[,4]))
  
  # calculate the total length of the scaffolds under consideration
  totLen = sum(as.numeric(b[,3]))
  
  # compute average depth of coverage per individual
  b$AveDepthInd = TotalDepthInd/totLen

  # compute average depth of coverage per scaffold
  b$AveScaffoldDepth = b$DepthCounts/b$EndPosition
  
  # Calculate the normalised depth per scaffold by dividing scaffold depth by total depth
  b$NormScaffoldDepth = b$AveScaffoldDepth / b$AveDepthInd

  # store all these subtables into a full data frame
  B <- rbind(B, b)
}

head(B)
```

Make an x axis
```{r}
scfIndex = as.data.frame(seq(1, 66, 1))
colnames(scfIndex) = c("scfIndex")
```



Visualize the normalized depths across scaffolds and per sample
```{r}
Z = ggplot(B[B$Sample != 'mwr_TG0457' & B$Sample != 'mwr_TG0459' & B$Sample != 'mwr_TG0464' & B$Sample != 'mwr_TGF08', ], aes(Scaffold, NormScaffoldDepth, colour=Sample)) +  geom_point(size=1, alpha=.6) + theme_bw() + guides(colour=F) + theme_bw()
  
#Z = Z + theme(axis.text.x=element_text(angle=60, hjust=1))
Z = Z + scale_y_continuous(limits=c(0,2)) + scale_x_discrete(labels=scfIndex$scfIndex, guide = guide_axis(n.dodge = 2))
Z = Z + ylab("Normalised depth of coverage")

Z
```



### Save scaffold-related plots as a supplementary figure

```{r}
pdf('WR_v4_FigS1.pdf', useDingbats = F)
 plot_grid(Y, Z, nrow=2, ncol=1, labels=c('A', 'B'), label_size = 10, align="v")
dev.off()
```


#### Who is the weirdo that shows a depth of coverage of 0 in 19/63 scaffolds? 
```{r}
ggplot(B[B$Sample == 'mwr_TG0463', ], aes(Scaffold, NormScaffoldDepth, colour=Sample)) +  geom_point(size=1, alpha=.6) + theme_bw()
```


It's mwr_TG0463 (SD1905.3)


##############################################################################

### Depth of coverage per sample across the final chosen scaffolds (n = 63)


Read in files
```{r}
# retrieve files
Depthfiles <- read.table('WR_56I_63scfs_13Mb.Depth', header=F)
```

```{r}
# initialize empty dataframe to collect the data and the new columns
A = NULL
```


Calculate depth per individual
```{r}
for (i in levels(Depthfiles$V5)) {
  
  # read in files
  a <- Depthfiles[Depthfiles$V5 == i, ]
  colnames(a) = c("Scaffold", "InitialPosition", "EndPosition", "DepthCounts", "Sample")
  
  # compute total depth of coverage per individual
  TotalDepthInd = sum(as.numeric(a[,4]))
  
  # calculate the total length of the scaffolds under consideration
  totLen = sum(as.numeric(a[,3]))
  
  # compute average depth of coverage per individual
  a$AveDepthInd = TotalDepthInd/totLen

  # compute average depth of coverage per scaffold
  a$AveScaffoldDepth = a$DepthCounts/a$EndPosition

  # store all these subtables into a full data frame
  A <- rbind(A, a)
}

head(A)

```


Plot the depth per scaffold per sample
```{r}
DS = ggplot(A, aes(Sample, AveScaffoldDepth, colour=Sample)) +  geom_violin(alpha=.6) + theme_bw() + guides(colour=F)
  
DS = DS + theme(axis.text.x=element_text(angle=90, size=3), axis.text.y=element_text(size=8), axis.title=element_text(size=8))
#DS = DS + scale_y_continuous(limits=c(0,2))
DS = DS + ylab("Depth of coverage")

DS
```



#### Plot per group

Subset the dataframe of depths into the 4 groups
```{r}
NWR1_names = c("mwr_TG0450", "mwr_TG0451", "mwr_TG0452", "mwr_TG0453", "mwr_TG0454", "mwr_TG0455", "mwr_TG0456", "mwr_TG0458", "mwr_TG0460", "mwr_TG0461", "mwr_TG0462", "mwr_TG0463", "mwr_TG0465", "mwr_TGF14", "mwr_TGF15", "mwr_TGF16")
NWR1 = A[A$Sample %in% NWR1_names,]

NWR2_names = c("wr_SRR5852741", "wr_SRR5852742", "wr_SRR5852743", "wr_SRR5852744", "wr_SRR5852745", "wr_SRR5852746", "wr_SRR5852747", "wr_SRR5852748", "wr_SRR5852749")
NWR2 = A[A$Sample %in% NWR2_names,]

SWR1_names = c("mwr_TG0449", "mwr_TGF01", "mwr_TGF02", "mwr_TGF03", "mwr_TGF04", "mwr_TGF05", "mwr_TGF06", "mwr_TGF07", "mwr_TGF09")
SWR1 = A[A$Sample %in% SWR1_names,]

SWR2_names = c("mwr_TGF10", "mwr_TGF11", "mwr_TGF12", "mwr_TGF13", "mwr_TGF17", "wr_P9109_102", "wr_P9109_103", "wr_P9109_104", "wr_P9109_105", "wr_P9109_106", "wr_P9109_107", "wr_P9109_108", "wr_P9109_109", "wr_P9109_110", "wr_SRR5852738", "wr_SRR5852739", "wr_SRR5852740", "wr_SRR5852750")
SWR2 = A[A$Sample %in% SWR2_names,]
```


Set the new labels
```{r}
NWR1_newID = c('CD-un.1','CD1912.1','CD1912.2','CD1911.1','CD1913.1','CD1912.3','CD1911.2','CD-un.2','SD1905.1','SD1905.2','UG1905.1','SD1905.3','SD1905.4','SD1914.1','UG1905.2','SD1905.5')

NWR2_newID = c('SD1972.1','SD1973.1','SS1954.1','SD1972.2','SD1974.1','SD1972.3','SD1963.1','UG1965.1','SD1972.4')
  
SWR1_newID = c('ZA1869.1','ZA1845.1','un1856.1','ZA-un.1','ZA1845.2','NA1856.1','NA1845.1','NA1857.1','ZA1842.1')
  
SWR2_newID = c('ZA1995.1','ZA2009.1','ZA2010.1','un2004.1','ZA1976.1','cap2009.1','cap2010.1','cap2008.1','cap2011.1','cap2011.2','cap2011.3','ZA2012.1','ZA1968.1','ZA1969.1','ZA1958.1','ZA1963.1','ZA1963.2','ZA1961.1')
```



Plot NWR1
```{r}
f = ggplot(NWR1, aes(Sample, AveScaffoldDepth)) 

f = f + 
  theme_bw() + theme(legend.position = "none", axis.text.x=element_text(angle=40, hjust=1, size=8)) +
  geom_violin(aes(Sample, AveScaffoldDepth), fill="gray40", colour="gold3", alpha=.4, lwd=.3) +
  scale_x_discrete(labels=as.character(NWR1_newID)) +
  scale_y_continuous(limits=c(0, 25)) +
  stat_summary(fun="median", geom="crossbar", width=.09, lwd=.2, colour="gray20") +
  labs(x="Sample name", y="Depth of coverage")

f
```


Plot NWR2
```{r}
g = ggplot(NWR2, aes(Sample, AveScaffoldDepth)) 

g = g + 
  theme_bw() + theme(legend.position = "none", axis.text.x=element_text(angle=40, hjust=1, size=8)) +
  geom_violin(aes(Sample, AveScaffoldDepth), fill="gray40", colour="darkorange", alpha=.4, lwd=.3) +
  scale_x_discrete(labels=as.character(NWR2_newID)) +
  scale_y_continuous(limits=c(0, 25)) +
  stat_summary(fun="median", geom="crossbar", width=.09, lwd=.2, colour="gray20") +
  labs(x="Sample name", y="Depth of coverage")

g
```

Plot SWR1
```{r}
h = ggplot(SWR1, aes(Sample, AveScaffoldDepth)) 

h = h + 
  theme_bw() + theme(legend.position = "none", axis.text.x=element_text(angle=40, hjust=1, size=8)) +
  geom_violin(aes(Sample, AveScaffoldDepth), fill="gray40", colour="royalblue4", alpha=.4, lwd=.3) +
  scale_x_discrete(labels=as.character(SWR1_newID)) +
  scale_y_continuous(limits=c(0, 25)) +
  stat_summary(fun="median", geom="crossbar", width=.09, lwd=.2, colour="gray20") +
  labs(x="Sample name", y="Depth of coverage")

h
```

Plot SWR2
```{r}
i = ggplot(SWR2, aes(Sample, AveScaffoldDepth)) 

i = i + 
  theme_bw() + theme(legend.position = "none", axis.text.x=element_text(angle=40, hjust=1, size=8)) +
  geom_violin(aes(Sample, AveScaffoldDepth), fill="gray40", colour="steelblue3", alpha=.4, lwd=.3) +
  scale_x_discrete(labels=as.character(SWR2_newID)) +
  scale_y_continuous(limits=c(0, 25)) +
  stat_summary(fun="median", geom="crossbar", width=.09, lwd=.2, colour="gray20") +
  labs(x="Sample name", y="Depth of coverage")

i

```

```{r}
figS2 = ggarrange(f + theme(axis.title.x=element_blank()) , g + theme(axis.text.y=element_blank(), axis.title=element_blank()), h, i + theme(axis.text.y=element_blank(), axis.title.y=element_blank()), align="v", labels=c("A", "B", "C", "D"), font.label=list(size=10))
figS2
```

```{r}
pdf('WR_v4_FigS2_v1.pdf', useDingbats = F)
figS2
dev.off()

```









