---
title: "WR Error Rates"
author: "Fátima S. Barreiro"
date: "October 8, 2020"
output: html_document
---

Load required libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(cowplot)
library(ggrepel)
library(reshape2)
library(dplyr)
library(tidyverse)
library(data.table)
require(scales)
library(ggpubr)
library(tidyr)
library(PairedData)
library(gridExtra)
```


Reading in file containing metadata and genomic diversity metrics 
```{r}
data = read.table("wr_P9109_110best_BRout.Est.txt", header=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```

Reading in file containing metadata and genomic diversity metrics 
```{r}
metadata = read.csv("WR_TableS1_v4.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```


```{r}
smalldata = data[1:52,]
IDs = as.data.frame(metadata$lab_ID)
colnames(IDs) = "ID"
smalldata = cbind(smalldata, IDs)
head(smalldata)
```


```{r}
smalldataM = reshape2::melt(smalldata)
head(smalldataM)
```

Plot
```{r}
E = ggplot(smalldataM, aes(ID, value, fill=variable, colour=variable)) + geom_bar(position = position_dodge(0.6), stat = "identity", width=0.4, size=0.2) + theme_bw() +
  theme(axis.text.x=element_text(angle=80, hjust=1), legend.key.size= unit(4, "mm"), legend.title=element_blank()) +
  ylab("Error rate") + xlab("Sample lab ID") +
  ggtitle("Error rates per sample")

E
```



```{r eval=FALSE, include=FALSE}
pdf("WR_ErrorRates.pdf", height=5, width=9, useDingbats = F)
E
dev.off()
```



### Correlations between error rates and estimates of heterozygosity

Parse the dataframe: add the heterozygosity estimates from the metadata
```{r}
head(smalldata)
HET = metadata$gw_het_GL2_3x
smalldataH = cbind(smalldata, HET)
head(smalldataH)
```


Calculate some basic correlation coefficients between the het estimate and the error rates separately
```{r}
num_cols = names(smalldataH)[sapply(smalldataH, is.numeric)] #get a list of numerical columns in the dataframe
length(num_cols)
cor(smalldataH$HET, smalldataH[num_cols]) #calculate correlations between all numerical correlations
```



Create a function to run and visualize linear regressions
```{r}

correlate_vars = function(dataframe, het_var, error_var, ID) {
  plot = ggscatter(dataframe, x = het_var, y = error_var, label = ID,
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson")
  return(plot)
  print(plot)
}

```


Use the function to visualise correlation plots of every error rate with het
```{r}
correlate_vars(smalldataH, "HET", "TtoC", "ID")
correlate_vars(smalldataH, "HET", "CtoT", "ID")
correlate_vars(smalldataH, "HET", "GtoA", "ID")
correlate_vars(smalldataH, "HET", "AtoG", "ID")
correlate_vars(smalldataH, "HET", "GtoT", "ID")
correlate_vars(smalldataH, "HET", "TtoG", "ID")
correlate_vars(smalldataH, "HET", "AtoT", "ID")
correlate_vars(smalldataH, "HET", "TtoA", "ID")
correlate_vars(smalldataH, "HET", "GtoC", "ID")
correlate_vars(smalldataH, "HET", "CtoG", "ID")
correlate_vars(smalldataH, "HET", "AtoC", "ID")
correlate_vars(smalldataH, "HET", "CtoA", "ID")

```

Significantly correlated with het:
G-to-T
T-to-G (marginally)
A-to-T
T-to-A
A-to-C (marginally)
C-to-A




### Correct the het estimates by the error rates that significantly correlate with het (i.e. the Tv error rates)


```{r}
head(smalldataH)

```


Calculate a composite error rate at Tv sites by summing up all error rates per row
```{r}
smalldataH %>% 
  dplyr::select(GtoT, TtoG, AtoT, TtoA, AtoC, CtoA) %>% 
  rowSums(na.rm=TRUE) -> smalldataH$ErTv

head(smalldataH)
```


Calculate the corrected het by operating with the composite error rate and the het estimate
```{r}
smalldataHc = smalldataH %>% 
  mutate(HET_corrected = HET-(HET*ErTv))

head(smalldataHc)
```

Save these newly calculated values as a file
```{r}
write.table(smalldataHc, file="GWhet_ErTv_corrected.csv", row.names=F, col.names=T, sep=",")
```


Visualise the new het estimates

```{r}
tinydataHc = smalldataHc %>%
  dplyr::select(ID, HET, HET_corrected)

tinydataHc = reshape2::melt(tinydataHc)
```



```{r}
P = ggplot(tinydataHc, aes(ID, value, colour=variable)) + geom_point() + theme_bw() +
  theme(axis.text.x=element_text(angle=80, hjust=1), legend.title = element_blank(), legend.position = "top") +
  scale_colour_manual(values=c("darkorange", "blue"), labels=c("gw_het_GL2_3x", "gw_het_GL2_3x_ErTvCorrected")) 
  
P
```


```{r}
pdf("WR_GWhet_ErTv.pdf", height=5, width=9, useDingbats = F)
P
dev.off()
```





















