---
title: "WR Error Rates"
author: "Fátima S. Barreiro"
date: "October 8, 2020"
output: html_document
---

Load required libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(cowplot)
library(ggrepel)
library(reshape2)
library(dplyr)
library(tidyverse)
library(data.table)
require(scales)
library(ggpubr)
library(tidyr)
library(PairedData)
library(gridExtra)
```


Reading in file containing metadata and genomic diversity metrics 
```{r}
data = read.table("wr_P9109_110best_BRout.Est.txt", header=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```

Reading in file containing metadata and genomic diversity metrics 
```{r}
metadata = read.csv("WR_TableS1_v4.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```


```{r}
smalldata = data[1:52,]
IDs = as.data.frame(metadata$lab_ID)
colnames(IDs) = "ID"
smalldata = cbind(smalldata, IDs)
head(smalldata)
```


```{r}
smalldataM = reshape2::melt(smalldata)
head(smalldataM)
```

Plot
```{r}
E = ggplot(smalldataM, aes(ID, value, fill=variable, colour=variable)) + geom_bar(position = position_dodge(0.6), stat = "identity", width=0.4, size=0.2) + theme_bw() +
  theme(axis.text.x=element_text(angle=80, hjust=1), legend.key.size= unit(4, "mm"), legend.title=element_blank()) +
  ylab("Error rate") + xlab("Sample lab ID") +
  ggtitle("Error rates per sample")

E
```



```{r eval=FALSE, include=FALSE}
pdf("WR_ErrorRates.pdf", height=5, width=9, useDingbats = F)
E
dev.off()
```



### Correlations between error rates and estimates of heterozygosity

Parse the dataframe: add the heterozygosity estimates from the metadata
```{r}
head(smalldata)
HET = metadata$gw_het_GL2_3x
smalldataH = cbind(smalldata, HET)
head(smalldataH)
```


Calculate some basic correlation coefficients between the het estimate and the error rates separately
```{r}
num_cols = names(smalldataH)[sapply(smalldataH, is.numeric)] #get a list of numerical columns in the dataframe
length(num_cols)
cor(smalldataH$HET, smalldataH[num_cols]) #calculate correlations between all numerical correlations
```



Create a function to run and visualize linear regressions
```{r}

correlate_vars = function(dataframe, het_var, error_var) {
  plot = ggscatter(dataframe, x = het_var, y = error_var, 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson")
  return(plot)
  print(plot)
}

```


Test the function with a variable
```{r}
correlate_vars(smalldataH, "HET", "T....C")
correlate_vars(smalldataH, "HET", "C....T")
correlate_vars(smalldataH, "HET", "G....A")
correlate_vars(smalldataH, "HET", "A....G")
correlate_vars(smalldataH, "HET", "G....T")
correlate_vars(smalldataH, "HET", "T....G")
correlate_vars(smalldataH, "HET", "A....T")
correlate_vars(smalldataH, "HET", "T....A")
correlate_vars(smalldataH, "HET", "G....C")
correlate_vars(smalldataH, "HET", "C....G")
correlate_vars(smalldataH, "HET", "A....C")
correlate_vars(smalldataH, "HET", "C....A")

```

Significantly correlated with het:
G-to-T
T-to-G (marginally)
A-to-T
T-to-A
A-to-C (marginally)
C-to-A


```{r}
for (i in seq_along(smalldataH)){
  if(is.numeric(i))
    correlate_vars(smalldataH, "HET", "i")
}

```































