---
title: "WR_structure_figures"
author: "Fátima S. Barreiro"
date: "February 16, 2020"
output: html_document
---

The code below is dedicated to visualizing the output of the structure analyses ran on the white rhino temporal genomics dataset. It includes PCA analysis, UMAP dimensionality reduction and an inactive line of code used for visualization of admixture results with pong. This corresponds to Figures 1B, 1C and S2 in the manuscript.

In order to thread chunks of code in R and in python, the R package 'reticulate' was required.



Load required libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggpubr)
library(tidyr)
library(cowplot)
```


## PCA with output from PCangsd v 0.973

```{r}
# prepare metadata
MD <- read.csv("WR_TableS1_v3.csv", h=T)

MD_unrelated = MD[MD$new_ID != 'CD-un.1' & MD$new_ID != 'un1856.1' & MD$new_ID != 'ZA1842.1', ]
```

```{r}
# prepare covariance matrix
D <-"WR.above13Mb_NoSexChr_NonMaskedRegions_nonRelated_tv.cov"
m <- as.matrix(read.table(D))

# run the eigen decomposition
e <- eigen(m)

# calculate the eigenvalues
eval = e$values/sum(e$values)
```


Select PCs to plot and bind them to metadata
```{r}
# PC1 and PC2
pcs12 <- e$vectors[,1:2]
evec12 <- cbind(pcs12, MD_unrelated)

# PC2 and PC3
pcs23 <- e$vectors[,2:3]
evec23 <- cbind(pcs23, MD_unrelated)
```


Visualize PCs with legend per subspecies, time window and country of origin
```{r}
## PC1 vs PC2
A = ggplot(evec12, aes(x=evec12$`1`, y=evec12$`2`))+
  
  # draw intercept lines
  geom_vline(xintercept=0, colour="gray", linetype = "longdash")+
  geom_hline(yintercept=0, colour="gray", linetype = "longdash")+
  
  # plot data points (2 legend alternatives)
  geom_point(aes(colour=evec12$time_window, shape=evec12$Country), alpha=.7, size=2) + 
  
  # assign colors and shapes
  scale_colour_manual(values=c("gold2", "darkorange", "royalblue4", "steelblue2")) +
  scale_shape_manual(values=c(7, 16, 10, 17, 18, 15, 9, 8)) +
  
  # labels, titles, background, etc.
  xlab(paste0("PC1 (", round(eval[1]*100, 2), "%)"))+
  ylab(paste0("PC2 (", round(eval[2]*100, 2), "%)"))+
  theme(panel.background = element_rect(fill='white', colour='black'), axis.text.x = element_text(colour='black'), axis.text.y = element_text(colour='black'), 
        strip.background = element_rect(size=.2, colour="black", fill ="#FFA5004D"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.title=element_text(size=0), legend.key=element_rect(fill="white", colour="white"), axis.title=element_text(size=8), legend.text=element_text(size=7),
        legend.key.size=unit(.3, "cm"))


## PC2 vs PC3
B = ggplot(evec23, aes(x=evec23$`1`, y=evec23$`2`))+
  
  # draw intercept lines
  geom_vline(xintercept=0, colour="gray", linetype = "longdash")+
  geom_hline(yintercept=0, colour="gray", linetype = "longdash")+
  
  # plot data points with legend per time window and country
  geom_point(aes(colour=evec23$time_window, shape=evec23$Country), alpha=.7, size=2) + 
  
  # assign colors and shapes to time windows and countries
  scale_colour_manual(values=c("gold2", "darkorange", "royalblue4", "steelblue2")) +
  scale_shape_manual(values=c(7, 16, 10, 17, 18, 15, 9, 8)) +
  
  # labels, titles, background, etc.
  xlab(paste0("PC2 (", round(eval[2]*100, 2), "%)"))+
  ylab(paste0("PC3 (", round(eval[3]*100, 2), "%)"))+
  theme(panel.background = element_rect(fill='white', colour='black'), axis.text.x = element_text(colour='black'), axis.text.y = element_text(colour='black'), 
        strip.background = element_rect(size=.2, colour="black", fill ="#FFA5004D"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.title=element_text(size=0), legend.key=element_rect(fill="white", colour="white"), axis.title=element_text(size=8), legend.text=element_text(size=7),
        legend.key.size=unit(.3, "cm"))


## place both PCA plots on same panel
plot_grid(A, B, labels=c('A', 'B'), ncol = 1, nrow = 2)
```



## Reduction of dimensionality of PC data with UMAP

### Prep the input for UMAP

```{r}
## Take the first 30 PCs in the covariance matrix
vectors_30 = e$vectors[,1:30]
values_30 = eval[1:30]

# Find out how much of the total variability is encompassed in these 30 PCs
sum(eval[1:30])

# Multiply each eigenvector by its corresponding eigenvalue
pcs_30 = sweep(vectors_30, 2, values_30, "*")

# Assign corresponding column names to the new dataframe
colnames(pcs_30) = c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10",
                     "PC11", "PC12", "PC13", "PC14", "PC15", "PC16", "PC17", "PC18", "PC19", "PC20", "PC21", 
                     "PC22", "PC23", "PC24", "PC25", "PC26", "PC27", "PC28", "PC29", "PC30")
```

```{r}
# write out as a table
write.table(pcs_30, "WR_NonRelated_30pcs.txt", sep="\t")
```



### Computing embeddings of PC data with UMAP

Install required modules
```{bash}
pip3 install numpy
pip3 install pandas
pip3 install umap
```


Load the required modules
```{python}
# import required packages and libraries
import numpy as np
import pandas as pd
import umap
```


Load the dataframe from PC data parsing above
```{python}
# import data and turn it into array
PCs = pd.read_csv("WR_NonRelated_30pcs.txt", sep='\t')
PCs.to_numpy()
```

Call the reducer function from UMAP
```{python}
reducer = umap.UMAP()
```

Reduce dimensionality with UMAP
```{python}
# train the model in umap with your data and reduce dimensionality of your dataset
embedding = reducer.fit_transform(PCs)

# the shape of the output should be n_samples x 2
embedding.shape 
```

Save the output for visualization
```{python}
# save output of umap as a text file to plot elsewhere
np.savetxt('WR_NonRelated_30pcs_embedding.txt', embedding, '%6.2f')
```


### Visualizing the embeddings from UMAP

Retrieve the output from UMAP
```{r}
# get embedding from umap
embeddings = read.table ("WR_NonRelated_30pcs_embedding.txt", h=F)
colnames(embeddings) = c('E1', 'E2')

emb = cbind(embeddings, MD_unrelated)
```

Visualize the 2-dimensional embeddings
```{r}
ggplot(emb, aes(E1, E2)) + 
  geom_point(aes(colour=emb$time_window, shape=emb$country), alpha=.7, size=2) + 
  
  # assign colors and shapes
  scale_colour_manual(values=c("gold2", "darkorange", "royalblue4", "steelblue2")) +
  scale_shape_manual(values=c(7, 16, 10, 17, 18, 15, 9, 8)) +
  
  xlab('UMAP1') + ylab('UMAP2') +
  theme(panel.background = element_rect(fill='white', colour='black'), axis.text.x = element_text(colour='black'), 
        axis.text.y = element_text(colour='black'), 
        strip.background = element_rect(size=.2, colour="black", fill ="#FFA5004D"), panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        legend.title=element_text(size=0), legend.key=element_rect(fill="white", colour="white"), 
        legend.key.size=unit(.3, 'cm'), legend.text=element_text(size=6),
        axis.title=element_text(size=6))
```



## Admixture analysis with output from ngsAdmix v 32 plot with pong

```{bash echo=TRUE}
#pong -m nonRelatedFilemap -i nonRelated_ind2pop_samples.txt -n nonRelated_poporder_final_newids.txt -l colourpalette_k6.txt
```

