---
title: "WR_v4_geneticLoad"
author: "Fátima S. Barreiro"
date: "November 13, 2020"
output: html_document
---


Load required libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(cowplot)
library(ggrepel)
library(reshape2)
library(dplyr)
library(tidyverse)
library(data.table)
require(scales)
library(ggpubr)
library(tidyr)
library(PairedData)
library(gridExtra)
```



Open error rates table
```{r}
snpeff = read.csv("snpeff_output.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```





## Plot number of homozygous derived sites per individual

```{r}
snpeff$group=factor(snpeff$group,levels=c("SWRpre Namibia","SWRpre South Africa","SWRpost South Africa", "SWRpost captive"))
```

Plot the number of hom and het derived sites
```{r}
ggplot(snpeff, aes(lab_ID, hom_derived)) + 
  geom_col(aes(group=cluster, fill=cluster)) +
  coord_flip()

ggplot(snpeff, aes(lab_ID, het_derived)) + 
  geom_col(aes(group=group, fill=group)) +
  coord_flip()
```



## Plot the ratio of lof/syn per individual

```{r}
homratio = ggplot(snpeff, aes(lab_ID, hom_ratio)) + theme_bw() +
  geom_col(aes(group=group, fill=group)) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(y="Homozygous derived LoF/synonymous ratio", x="Sample ID") +
  theme(legend.title=element_blank())

homratio
```




## Plotting per time window
```{r}
snpeff$time_window=factor(snpeff$time_window,levels=c("SWRpre", "SWRpost"))
```


```{r}
# the hom ratio
homratio1 = ggplot(snpeff, aes(time_window, hom_ratio)) + theme_bw() +
  
  # plot distribution and points
  geom_violin(aes(group=time_window, colour=time_window), fill="gray70", lwd=.5, alpha=.4) +
  geom_point(aes(time_window, hom_ratio, group=time_window, colour=time_window), shape=17, cex=2, alpha=.7) +
  
  # add colour scale
  scale_color_manual(values= (c("royalblue4", "steelblue3"))) +
  
  # format axes and labels
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1), limits=c(0.013, 0.027), breaks=seq(0.014, 0.026, 0.004)) +
  labs(y="LoF/synonymous ratio at homozygous derived sites", x=NULL) +
  theme(legend.position = "none", axis.text.x=element_text(size=12)) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='gray30', width=.12, lwd=.4) +
  
  # run comparison with wilcox test
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"), label.y = c(0.026, 0.027), colour="gray20") +
  
  # add labels to SWRpre
  geom_text_repel(data=subset(snpeff, time_window!="SWRpost"), aes(label=sample), colour="grey20", size=3, point.padding=1, box.padding=1, direction="x", segment.size=.2)

homratio1


```



## Plotting per time window and Country

```{r}
snpeff$group=factor(snpeff$group,levels=c("SWRpre Namibia","SWRpre South Africa","SWRpost South Africa", "SWRpost captive"))
```


```{r}
# the hom ratio
homratio2 = ggplot(snpeff, aes(group, hom_ratio)) + theme_bw() +
  geom_boxplot(aes(group=group, colour=group), fill="gray70", lwd=.5, alpha=.4) +
  geom_jitter(aes(group, hom_ratio, group=group, colour=group), shape=17, position=position_jitter(0.11), cex=2, alpha=.7) +
  
  scale_color_manual(values= (c("royalblue4", "#4169e1", "#0077be", "steelblue3"))) +
  
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(y="LoF/synonymous ratio at homozygous derived sites", x=NULL) +
  theme(legend.position = "none", axis.text.x=element_text(angle=20, hjust=1)) +
  
  geom_text_repel(data=snpeff, aes(label=sample), segment.size=.15, size=2)

homratio2


```



## Plotting per time window and cluster

```{r}
snpeff$cluster=factor(snpeff$cluster,levels=c("SWRpre West", "SWRpre East" ,"SWRpost wild", "SWRpost captive"))
```

```{r}
# the hom ratio
homratio3 = ggplot(snpeff, aes(cluster, hom_ratio)) + theme_bw() +
  geom_boxplot(aes(group=cluster, colour=cluster), fill="gray70", lwd=.5, alpha=.4) +
  geom_jitter(aes(cluster, hom_ratio, group=cluster, colour=cluster), shape=17, position=position_jitter(0.11), cex=2, alpha=.7) +
  
  scale_color_manual(values= (c("royalblue4", "#4169e1", "#0077be", "steelblue3"))) +
  
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1), limits=c(0.013, 0.027), breaks=seq(0.014, 0.026, 0.004)) +
  
  labs(y="LoF/synonymous ratio at homozygous derived sites", x=NULL) +
  theme(legend.position = "none", axis.text.x=element_text(angle=30, hjust=1, size=12)) +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("SWRpre West", "SWRpre East")), method.args = list(alternative = "greater"), label.y = c(0.026)) +
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("SWRpost wild", "SWRpost captive")), method.args = list(alternative = "two.sided"), label.y = c(0.024)) 

  #geom_text_repel(data=snpeff, aes(label=sample), segment.size=.15, size=2)

homratio3


```


## Make composite figure for manuscript

```{r}
homratio_fig = ggarrange(homratio1, 
          homratio3 + theme(axis.text.y = element_blank(),
                     axis.ticks.y = element_blank(),
                     axis.title.y = element_blank()), 
          nrow=1, ncol=2, align="h", widths=c(1.2, 1), labels=c("A", "B"))

homratio_fig
```



```{r}
pdf("WR_v4_Fig6_GeneticLoad.pdf", height=5, useDingbats = F)
homratio_fig
dev.off()
```


## Comparison of genomic metrics in SWRpre of interest

Open metadata and genomic metrics table
```{r}
md = read.csv("WR_TableS1_v5.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```

Subset the table to keep only the samples of interest
```{r}
swrpre
```

```{r}
swrpre = subset(md, sample_name == "NA1845.1" | sample_name == "NA1857.1" |  sample_name == "ZA-un.1" | sample_name == "NA1856.1" | sample_name == "ZA1869.1")
```


Create a cluster variable to separate the West and the East samples
```{r}
swrpre$Cluster = c("East", "West", "East", "West", "West")
```


Order the levels of the Cluster variable for plotting
```{r}
swrpre$sample_name = with(swrpre, factor(sample_name, levels=sample_name[order(Cluster)]))
```


### GW Het in SWRpre samples of interest

```{r}
swrpreH = ggplot(swrpre, aes(sample_name, gw_het_GL2_5x_ErTv, colour=Cluster, fill=Cluster)) + theme_bw() +
  geom_col(alpha=.9) +
  labs(x="Sample name", y=expression("Genome-wide heterozygosity")) +
  scale_fill_manual(values=c("royalblue4", "#4169e1")) +
  scale_colour_manual(values=c("royalblue4", "#4169e1")) +
  coord_flip()

swrpreH
```


### Froh in SWRpre samples of interest

```{r}
swrpreF = ggplot(swrpre, aes(sample_name, Froh, fill=Cluster, colour=Cluster)) + theme_bw() +
  geom_col(alpha=.9) +
  labs(x="Sample name", y=expression("F"["RoH"])) +
  scale_fill_manual(values=c("royalblue4", "#4169e1")) +
  scale_colour_manual(values=c("royalblue4", "#4169e1")) +
  coord_flip() 

swrpreF
```




### The genetic load in SWRpre of interest

Plot the genetic load in the SWRpre of interest

FINISH


```{r}
swrpreG = ggplot(data=subset(snpeff, sample == "NA1845.1" | sample == "NA1857.1" |  sample == "ZA-un.1" | sample == "NA1856.1" | sample == "ZA1869.1"), 
                 aes(sample, hom_ratio), fill=cluster, colour=cluster, group=cluster) + theme_bw() +
  geom_col(alpha=.9) +
  labs(x="Sample name", y="LoF/synonymous ratio at homozygous derived sites") +
  scale_fill_manual(values=c("royalblue4", "#4169e1")) +
  scale_colour_manual(values=c("royalblue4", "#4169e1")) +
  coord_flip() 


swrpreG
```






```{r}
swrpreG = ggplot(snpeff, aes(sample_name, hom_ratio)) + theme_bw() +
```

















```{r}
figS9 = ggarrange(swrpreH + guides(colour=F, fill=F), swrpreF + theme(axis.text.y=element_blank(), axis.title.y=element_blank(), axis.ticks.y=element_blank()), nrow=1, ncol=2, align="h")
figS9
```


```{r}
pdf("WR_v4_FigS9_SWRpre.pdf", height=5, useDingbats = F)
homratio_fig
dev.off()
```


















