---
title: "WR_v4 Het tests"
author: "Fátima S. Barreiro"
date: "October 20, 2020"
output: html_document
---


Load required libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(cowplot)
library(ggrepel)
library(reshape2)
library(dplyr)
library(tidyverse)
library(data.table)
require(scales)
library(ggpubr)
library(tidyr)
library(PairedData)
library(gridExtra)
```



Open error rates table
```{r}
error = read.csv("GWhet_errors_ErTv.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```


Open metadata and genomic metrics table
```{r}
data = read.csv("WR_TableS1_v5.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```

Make a dataframe with relevant columns
```{r}
HET3x = as.data.frame(data$gw_het_GL2_3x)
HET5x = as.data.frame(data$gw_het_GL2_5x)
ErTv = as.data.frame(error$ErTv)
ID1 = as.data.frame(data$lab_ID)
ID2 = as.data.frame(error$ID)
```

```{r}
correct = cbind(ID1, ID2, ErTv, HET3x, HET5x)
colnames(correct) = c("ID1", "ID2", "ErTv", "HET3x", "HET5x")
head(correct)
```


Calculate estimates of het corrected by ErTv

```{r}
correct = correct %>%
  mutate(HET3x_corrected = HET3x-(HET3x*ErTv), HET5x_corrected = HET5x-(HET5x*ErTv))

head(correct)

```


Prepare for plotting

```{r}
correct2plot = correct %>%
  dplyr::select(ID1, HET3x, HET5x, HET3x_corrected, HET5x_corrected)
  
correct2plot = reshape2::melt(correct2plot)
```


```{r}
ggplot(correct2plot, aes(ID1, value, colour=variable)) + geom_point() + theme_bw() +
  theme(axis.text.x=element_text(angle=80, hjust=1), legend.title = element_blank(), legend.position = "top") 
  #scale_colour_manual(values=c("darkorange", "blue"), labels=c("gw_het_GL2_3x_ErTv", "gw_het_GL2_5x_ErTv"))
```



## Estimates of Het across groups and group comparisons


### With GL2 <3x estimate excluding the 4 samples of average depth of coverage <5x

```{r}
data$time_window=factor(data$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```



```{r}
# plot het values per time window
gw_het_3x = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(time_window, gw_het_GL2_3x_ErTv)) + 
  theme_bw() + theme(legend.position = "none", axis.text=element_text(size=7, colour='gray10'), axis.title.y=element_text(size=7), 
                     panel.grid.major=(element_line(size=.12, colour='gray80')))
  
  
# plot violin per group
gw_het_3x = gw_het +  geom_violin(aes(colour=time_window, fill=time_window), lwd=.3, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(colour='gray40', shape=16, position=position_jitter(0.11), cex=.6, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='black', width=.12, lwd=.3) +
  
  stat_summary(fun=mean, fun.min = function (gw_het_GL2_3x_ErTv) mean(gw_het_GL2_3x_ErTv) - sd(gw_het_GL2_3x_ErTv), 
               fun.max = function(gw_het_GL2_3x_ErTv) mean(gw_het_GL2_3x_ErTv) + sd(gw_het_GL2_3x_ErTv), geom = "pointrange" , size=.12, stroke=.4, col="black") +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"), label.y = c(0.00024, 0.00020), size=2.5) +
  
  # embellish
  scale_y_continuous(limits=c(0.00007, 0.00025),labels = function(x) format(x, scientific = FALSE)) +
  
  
  labs(x=NULL, y="Genome-wide heterozygosity (tv)") + geom_text_repel(data=data, aes(label=lab_ID), segment.size=.15, size=2)

```



### With GL2 <5x estimate excluding the 4 samples of average depth of coverage <5x

First set the order of the time windows as they should appear on the plot
```{r}
data$time_window=factor(data$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```

```{r}
# plot het values per time window
gw_het = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(time_window, gw_het_GL2_5x_ErTv)) + 
  theme_bw() + theme(legend.position = "none") + labs(x=NULL, y="Genome-wide heterozygosity") +
  scale_y_continuous(labels = function(x) format(x, scientific = F), limits=c(0.00008, 0.00027))
  
  
# plot violin per group
gw_het = gw_het +  
  geom_violin(aes(colour=time_window), fill="gray70", lwd=.5, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(aes(colour=time_window), shape=17, position=position_jitter(0.11), cex=2, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='gray30', width=.2, lwd=.4)
  

gw_het = gw_het + stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"), label.y = c(0.00026, 0.00025), size=3, colour="gray20")   
#+ #geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=sample_name), segment.size=.15, size=2)

gw_het
```








#############################################################################

## RoH length and FRoH 

After parsing the output of BCFtools RoH in the server, the resulting values were copy-pasted to the metadata/data spreadsheet used here.
Now visualize and compare temporal groups in terms of FRoH and RoH length.

First set the order of the time windows as they should appear on the plot
```{r}
data$time_window=factor(data$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```



#### Plot FRoH per individual in time-window groups
```{r}
# plot het values per time window
fRoh = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TGF02' & lab_ID!='mwr_TGF04' & lab_ID!='mwr_TG0452' & lab_ID!='mwr_TG0461'),
       aes(time_window, Froh_10scf)) +
  
  # format scale in y axis
  scale_y_continuous(limits=c(0, 0.55), breaks=seq(0, 0.55, 0.1), labels=scales::number_format(accuracy = 0.01)) +

# set up the panel
theme_bw() + theme(legend.position = "none") + labs(x=NULL, y=expression("F"["RoH"]))

  # plot violin per group
fRoh = fRoh +  
  geom_violin(aes(colour=time_window), fill="gray70", lwd=.5, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(aes(colour=time_window), shape=17, position=position_jitter(0.11), cex=2, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='gray30', width=.2, lwd=.4)
  
  fRoh = fRoh + stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "less"), label.y=c(0.5, 0.45), size=3, colour="gray20") 
  #+
  
  #geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=sample_name), segment.size=.15, size=2)
  

fRoh
```



#### Plot average RoH length per individual in the 10 largest scaffolds


Visualize average RoH length in the 10 largest scaffolds
```{r}
aveRoh = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TGF02' & lab_ID!='mwr_TGF04' & lab_ID!='mwr_TG0452' & lab_ID!='mwr_TG0461'),
       aes(time_window, ave_RoH_len)) +
  
  scale_y_continuous(limits=c(50000, 840000), labels = function(x) format(x/1000000, digits=3, nsmall=2), breaks=seq(100000, 800000, 100000)) +

# set up the panel
theme_bw() + theme(legend.position = "none") + labs(x=NULL, y="Average RoH length (Mbp)")

# plot violin per group
aveRoh = aveRoh +  
  geom_boxplot(aes(colour=time_window), fill="gray70", lwd=.5, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(aes(colour=time_window), shape=17, position=position_jitter(0.11), cex=2, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) 
  
# perform wilcox tests to compare groups
aveRoh = aveRoh + stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "less"), label.y=c(600000, 800000), size=3, colour="gray20") #+
  
  #geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=sample_name), segment.size=.15, size=2)

aveRoh
  

```




#### Visualize maximum RoH length in the 10 largest scaffolds

Use the function written before to customize the y axis here too

```{r}
maxRoh = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TGF02' & lab_ID!='mwr_TGF04' & lab_ID!='mwr_TG0452' & lab_ID!='mwr_TG0461'),
       aes(time_window, max_RoH_len)) +
  
  # customize the y axis scale
  scale_y_continuous(limits=c(100000, 16000000), labels = function(x) format(x/1000000, digits=3, nsmall=2), breaks=seq(1000000, 16000000, 2000000)) +

# set up the panel
theme_bw() + theme(legend.position = "none") + labs(x=NULL, y="Maximum RoH length (Mbp)")


maxRoh = maxRoh +  
  # create boxplots
  geom_boxplot(aes(colour=time_window), fill="gray60", lwd=.5, alpha=.3) +
  
  # add scatter points per group
  geom_jitter(aes(colour=time_window), shape=17, position=position_jitter(0.11), cex=2, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # run the wilcox tests to compare groups
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "less"), label.y=c(14000000, 15000000), size=3, colour="gray20") #+
  
  #geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=sample_name), segment.size=.15, size=2)
  
maxRoh

```




### Make composite figure for genomic erosion stats

```{r}
erosion = ggarrange(gw_het + theme(axis.text.x=element_blank()), 
                    fRoh + theme(axis.text.x=element_blank()), 
                    aveRoh + theme(axis.text.x=element_text(size=10)), 
                    maxRoh + theme(axis.text.x=element_text(size=10)), 
                    nrow=2, ncol=2, align="v", labels=c( "A", "B" , "C", "D"))

erosion
```


```{r}
pdf("WR_v4_GenomicErosionFig.pdf", useDingbats = F)
erosion
dev.off()
```

















Visualize min RoH length in the 10 largest scaffolds

```{r}
ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TGF02' & lab_ID!='mwr_TGF04' & lab_ID!='mwr_TG0452' & lab_ID!='mwr_TG0461'),
       aes(time_window, min_RoH_len)) +


# set up the panel
theme_bw() + theme(legend.position = "none", axis.text=element_text(size=7, colour='gray10'), axis.title.y=element_text(size=7), 
                     panel.grid.major=(element_line(size=.12, colour='gray80'))) +

# plot violin per group
geom_violin(aes(colour=time_window, fill=time_window), lwd=.3, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(colour='gray40', shape=16, position=position_jitter(0.11), cex=.6, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='black', width=.12, lwd=.3) +
  
  stat_summary(fun=mean, fun.min = function (min_RoH_len) mean(min_RoH_len) - sd(min_RoH_len), 
               fun.max = function(min_RoH_len) mean(min_RoH_len) + sd(min_RoH_len), geom = "pointrange" , size=.12, stroke=.4, col="black") +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "less")) +
  
  geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=sample_name), segment.size=.15, size=2)
  

```



## Extra analyses


### Results of ngsF (inbreeding coefficients per group)

```{r}
data$time_window=factor(data$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```

```{r}
# plot het values per time window
indFplot = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TGF02' & lab_ID!='mwr_TGF09'), 
              aes(time_window, F_ngsF)) + 
  theme_bw() + theme(legend.position = "none", axis.text=element_text(size=7, colour='gray10'), axis.title.y=element_text(size=7), 
                     panel.grid.major=(element_line(size=.12, colour='gray80')))
  
  
# plot violin per group
indFplot = indFplot +  geom_violin(aes(colour=time_window, fill=time_window), lwd=.3, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(colour='gray40', shape=16, position=position_jitter(0.11), cex=.6, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='black', width=.12, lwd=.3) +
  
  stat_summary(fun=mean, fun.min = function (F_ngsF) mean(F_ngsF) - sd(F_ngsF), 
               fun.max = function(F_ngsF) mean(F_ngsF) + sd(F_ngsF), geom = "pointrange" , size=.12, stroke=.4, col="black") +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"), label.y = c(0.00027, 0.00020), size=2.5) #+
  
  # embellish
  #scale_y_continuous(limits=c(0.00007, 0.00030),labels = function(x) format(x, scientific = FALSE)) +
  
  
#  labs(x=NULL, y="Inbreeding coefficient") + geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TGF02' & lab_ID!='mwr_TGF09'), aes(label=lab_ID), segment.size=.15, size=2)

indFplot
```




### GW het after downsampling at 5x all samples

```{r}
data$time_window=factor(data$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```

```{r}
# plot het values per time window
gw_hetDS = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(time_window, GW_het_DS_corrected)) + 
  theme_bw() + theme(legend.position = "none") + labs(x=NULL, y="Genome-wide heterozygosity (tv)") +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
  
  
# plot violin per group
gw_hetDS = gw_hetDS +  
  geom_violin(aes(colour=time_window), fill="gray70", lwd=.5, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(aes(colour=time_window), shape=17, position=position_jitter(0.11), cex=2, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='gray30', width=.16, lwd=.4)
  

gw_hetDS = gw_hetDS + stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"))   + geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=sample_name), segment.size=.15, size=2)

gw_hetDS
```



### Het estimates in regions annotated as exons

```{r}
# plot het values per time window
exon_het = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(time_window, exon_het_GL2_5x_ErTv)) + 
  theme_bw() + theme(legend.position = "none") + labs(x=NULL, y="Exon heterozygosity (tv)") +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
  
  
# plot violin per group
exon_het = exon_het +  
  geom_violin(aes(colour=time_window), fill="gray70", lwd=.5, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(aes(colour=time_window), shape=17, position=position_jitter(0.11), cex=2, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='gray30', width=.16, lwd=.4)
  
exon_het = exon_het + stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"))

exon_het
```












