---
title: "WR_v4_PCAs"
author: "Fátima S. Barreiro"
date: "October 29, 2020"
output: html_document
---

Load required libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggpubr)
library(tidyr)
library(cowplot)
library(ggrepel)
```


```{r}
# load metadata
md = read.csv("WR_TableS1_v5.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")

# make a subset where related inds (1 from each pair) has been removed
md_unrel = md[md$sample_name != 'CD-un.1' & md$sample_name != 'un1856.1' & md$sample_name != 'ZA1842.1', ]
```


## PCA of all 4 groups, entire unrelated dataset (49 samples)

Open the covariance matrix and do the eigen decomposition into vectors and values
```{r}
# prepare covariance matrix
D <-"WR.49I.above13Mb_NoSexChr_NonMaskedRegions_GL2_unrelated.tv.cov"
m <- as.matrix(read.table(D))

# run the eigen decomposition
e <- eigen(m)

# calculate the eigenvalues
eval = e$values/sum(e$values)
```


Select PCs to plot and bind them to metadata
```{r}
# PC1 and PC2
pcs <- as.data.frame(e$vectors[,1:3])
colnames(pcs) = c("PC1", "PC2", "PC3")

# Extract relevant columns from metadata for plotting
mini_md = as.data.frame(md_unrel %>% select(time_window, country))

evec <- cbind(mini_md, pcs)
head(evec)
```

Make a composite column for the legend
```{r}
evec$window_country <- paste(evec$time_window, evec$country, sep = "-")
head(evec)
```


Set the right order of factors for the legend
```{r}
evec$time_window=factor(evec$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```


Visualize PCs with legend per subspecies, time window and country of origin
```{r}
## PC1 vs PC2
A = ggplot(evec, aes(x=PC1, y=PC2)) +
  
  # draw intercept lines
  geom_vline(xintercept=0, colour="gray", linetype = "longdash")+
  geom_hline(yintercept=0, colour="gray", linetype = "longdash")+
  
  # plot data points (2 legend alternatives)
  geom_point(aes(colour = time_window, shape = time_window), alpha=.7, size=4) + 
  
  # assign colors and shapes
  scale_colour_manual(values=c("gold3", "darkorange", "royalblue4", "steelblue3")) +
  scale_shape_manual(values=c(15, 16, 17, 8)) +
  
  # labels, titles, background, etc.
  xlab(paste0("PC1 (", round(eval[1]*100, 2), "%)"))+
  ylab(paste0("PC2 (", round(eval[2]*100, 2), "%)"))+
  
  theme(panel.background = element_rect(fill='white', colour='black'), axis.text.x = element_text(colour='black'), axis.text.y = element_text(colour='black'), 
        strip.background = element_rect(size=.2, colour="black", fill ="#FFA5004D"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.title=element_text(size=0), legend.key=element_rect(fill="white", colour="white"), axis.title=element_text(size=10), legend.text=element_text(size=8), legend.key.size=unit(.8, "cm"),
        legend.position = "top", legend.margin=margin(0,0,0,0), legend.box.margin=margin(0,-10,-10,-10), plot.margin = unit(c(0,0,0,0.2), "lines"))

A
```


```{r}
## PC2 vs PC3
B = ggplot(evec, aes(x=PC2, y=PC3)) +
  
  # draw intercept lines
  geom_vline(xintercept=0, colour="gray", linetype = "longdash")+
  geom_hline(yintercept=0, colour="gray", linetype = "longdash")+
  
  # plot data points (2 legend alternatives)
  geom_point(aes(colour = time_window, shape = time_window), alpha=.7, size=4) + 
  
  # assign colors and shapes
  scale_colour_manual(values=c("gold3", "darkorange", "royalblue4", "steelblue3")) +
  scale_shape_manual(values=c(15, 16, 17, 8)) +
  
  # labels, titles, background, etc.
  xlab(paste0("PC2 (", round(eval[2]*100, 2), "%)"))+
  ylab(paste0("PC3 (", round(eval[3]*100, 2), "%)"))+
  
  theme(panel.background = element_rect(fill='white', colour='black'), axis.text.x = element_text(colour='black'), axis.text.y = element_text(colour='black'), 
        strip.background = element_rect(size=.2, colour="black", fill ="#FFA5004D"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.title=element_text(size=0), legend.key=element_rect(fill="white", colour="white"), axis.title=element_text(size=10), legend.text=element_text(size=10), legend.key.size=unit(.8, "cm"), 
        legend.position = "top", legend.margin=margin(0,0,0,0), legend.box.margin=margin(0,-10,-10,-10), plot.margin=unit(c(0,0,0,1), "cm"))

B
```






########################################################################################################

## PCA's per subspecies


### NWR

Open the covariance matrices of NWR or SWR and do the eigen decomposition into vectors and values
```{r}
# prepare covariance matrix
NWR <-"NWR.24I.above13Mb_NoSexChr_NonMaskedRegions_GL2_unrelated.tv.cov"
mNWR <- as.matrix(read.table(NWR))

# run the eigen decomposition
eNWR <- eigen(mNWR)

# calculate the eigenvalues
evalNWR = eNWR$values/sum(eNWR$values)
```


Prepare the metadata
```{r}
mdNWR = md_unrel[md_unrel$subspecies == "cottoni", ]
```


Select PCs to plot and bind them to metadata
```{r}
# PC1 and PC2
pcsNWR <- as.data.frame(eNWR$vectors[,1:3])
colnames(pcsNWR) = c("PC1", "PC2", "PC3")

# Extract relevant columns from metadata for plotting
mini_mdNWR = as.data.frame(mdNWR %>% select(sample_name, lab_ID, time_window, country))

evecNWR <- cbind(mini_mdNWR, pcsNWR)
head(evecNWR)
```



Visualize PCs with legend per subspecies, time window and country of origin

```{r}
evecNWR$time_window=factor(evecNWR$time_window,levels=c("NWRpre","NWRpost"))
```


```{r}
## PC1 vs PC2
nwr = ggplot(evecNWR, aes(x=PC1, y=PC2)) +
  
  # draw intercept lines
  geom_vline(xintercept=0, colour="gray", linetype = "longdash")+
  geom_hline(yintercept=0, colour="gray", linetype = "longdash")+
  
  # plot data points (2 legend alternatives)
  geom_point(aes(colour = time_window, shape = country), alpha=.7, size=3) + 
  
  # assign colors and shapes
  scale_colour_manual(values=c("gold3", "darkorange")) +
  scale_shape_manual(values=c(15, 18, 17, 8)) +
  
  
  # labels, titles, background, etc.
  xlab(paste0("PC1 (", round(evalNWR[1]*100, 2), "%)"))+
  ylab(paste0("PC2 (", round(evalNWR[2]*100, 2), "%)"))+
  
  theme(panel.background = element_rect(fill='white', colour='black'), axis.text.x = element_text(colour='black'), axis.text.y = element_text(colour='black'), 
        strip.background = element_rect(size=.2, colour="black", fill ="#FFA5004D"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.title=element_text(size=0), legend.key=element_rect(fill="white", colour="white"), axis.title=element_text(size=10), legend.text=element_text(size=7), legend.key.size=unit(.4, "cm"),
        legend.box.margin=margin(-10,-10,-10,-10), plot.margin = unit(c(1,0.5,0,0), "lines")) +
  
  guides(colour = guide_legend(order = 1), shape = guide_legend(order = 2)) #+ geom_text_repel(data=evecNWR, aes(label=sample_name), segment.size=.15, size=2)

nwr
```



### SWR

Open the covariance matrices of NWR or SWR and do the eigen decomposition into vectors and values
```{r}
# prepare covariance matrix
SWR <-"SWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2_unrelated.tv.cov"
mSWR <- as.matrix(read.table(SWR))

# run the eigen decomposition
eSWR <- eigen(mSWR)

# calculate the eigenvalues
evalSWR = eSWR$values/sum(eSWR$values)
```


Prepare the metadata
```{r}
mdSWR = md_unrel[md_unrel$subspecies == "simum", ]
```


Select PCs to plot and bind them to metadata
```{r}
# PC1 and PC2
pcsSWR <- as.data.frame(eSWR$vectors[,1:3])
colnames(pcsSWR) = c("PC1", "PC2", "PC3")

# Extract relevant columns from metadata for plotting
mini_mdSWR = as.data.frame(mdSWR %>% select(sample_name, lab_ID, time_window, country))

evecSWR <- cbind(mini_mdSWR, pcsSWR)
head(evecSWR)
```


```{r}
evecSWR$time_window=factor(evecSWR$time_window,levels=c("SWRpre","SWRpost"))
```


```{r}
## PC1 vs PC2
swr = ggplot(evecSWR, aes(x=PC1, y=PC2)) +
  
  # draw intercept lines
  geom_vline(xintercept=0, colour="gray", linetype = "longdash")+
  geom_hline(yintercept=0, colour="gray", linetype = "longdash")+
  
  # plot data points (2 legend alternatives)
  geom_point(aes(colour = time_window, shape = country), alpha=.7, size=3) + 
  
  # assign colors and shapes
  scale_colour_manual(values=c("royalblue4", "steelblue3")) +
  scale_shape_manual(values=c(15, 18, 17, 8)) +
  
  # labels, titles, background, etc.
  xlab(paste0("PC1 (", round(evalSWR[1]*100, 2), "%)"))+
  ylab(paste0("PC2 (", round(evalSWR[2]*100, 2), "%)"))+
  
  theme(panel.background = element_rect(fill='white', colour='black'), axis.text.x = element_text(colour='black'), axis.text.y = element_text(colour='black'), 
        strip.background = element_rect(size=.2, colour="black", fill ="#FFA5004D"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.title=element_text(size=0), legend.key=element_rect(fill="white", colour="white"), axis.title=element_text(size=10), legend.text=element_text(size=7), legend.key.size=unit(.4, "cm"), 
        legend.box.margin=margin(-10,-10,-10,-10), plot.margin = unit(c(1,0.5,0,0), "lines")) +
  
  guides(colour = guide_legend(order = 1), shape = guide_legend(order = 2)) +
  geom_text_repel(data=subset(evecSWR, sample_name == "ZA1845.2"), aes(label=sample_name), segment.size=0, size=3, colour="gray20", ylim=c(-0.32, -0.1), xlim=c(-0.25, -0.05))
  
   

swr
```



```{r}
subsp = ggarrange(nwr, swr, nrow = 2, ncol = 1, labels=c("B", "C"))
```


```{r}
PCAps = ggarrange(A, subsp, nrow=1, ncol=2, labels=c("A", NULL), align = c("v"))
PCAps
```



```{r}
pdf('WR_v4_PCAs.pdf', width=9, useDingbats = F)
PCAps
dev.off()
```









