---
title: "WR_v4 Het tests"
author: "Fátima S. Barreiro"
date: "October 20, 2020"
output: html_document
---


Load required libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(cowplot)
library(ggrepel)
library(reshape2)
library(dplyr)
library(tidyverse)
library(data.table)
require(scales)
library(ggpubr)
library(tidyr)
library(PairedData)
library(gridExtra)
```



Open error rates table
```{r}
error = read.csv("GWhet_errors_ErTv.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```


Open metadata and genomic metrics table
```{r}
data = read.csv("WR_TableS1_v5.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```

Make a dataframe with relevant columns
```{r}
HET3x = as.data.frame(data$gw_het_GL2_3x)
HET5x = as.data.frame(data$gw_het_GL2_5x)
ErTv = as.data.frame(error$ErTv)
ID1 = as.data.frame(data$lab_ID)
ID2 = as.data.frame(error$ID)
```

```{r}
correct = cbind(ID1, ID2, ErTv, HET3x, HET5x)
colnames(correct) = c("ID1", "ID2", "ErTv", "HET3x", "HET5x")
head(correct)
```


Calculate estimates of het corrected by ErTv

```{r}
correct = correct %>%
  mutate(HET3x_corrected = HET3x-(HET3x*ErTv), HET5x_corrected = HET5x-(HET5x*ErTv))

head(correct)

```


Prepare for plotting

```{r}
correct2plot = correct %>%
  dplyr::select(ID1, HET3x, HET5x, HET3x_corrected, HET5x_corrected)
  
correct2plot = reshape2::melt(correct2plot)
```


```{r}
ggplot(correct2plot, aes(ID1, value, colour=variable)) + geom_point() + theme_bw() +
  theme(axis.text.x=element_text(angle=80, hjust=1), legend.title = element_blank(), legend.position = "top") 
  #scale_colour_manual(values=c("darkorange", "blue"), labels=c("gw_het_GL2_3x_ErTv", "gw_het_GL2_5x_ErTv"))
```



## Estimates of Het across groups and group comparisons


### With GL2 <3x estimate excluding the 4 samples of average depth of coverage <5x

```{r}
data$time_window=factor(data$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```



```{r}
# plot het values per time window
gw_het = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(time_window, gw_het_GL2_3x_ErTv)) + 
  theme_bw() + theme(legend.position = "none", axis.text=element_text(size=7, colour='gray10'), axis.title.y=element_text(size=7), 
                     panel.grid.major=(element_line(size=.12, colour='gray80')))
  
  
# plot violin per group
gw_het = gw_het +  geom_violin(aes(colour=time_window, fill=time_window), lwd=.3, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(colour='gray40', shape=16, position=position_jitter(0.11), cex=.6, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='black', width=.12, lwd=.3) +
  
  stat_summary(fun=mean, fun.min = function (gw_het_GL2_3x_ErTv) mean(gw_het_GL2_3x_ErTv) - sd(gw_het_GL2_3x_ErTv), 
               fun.max = function(gw_het_GL2_3x_ErTv) mean(gw_het_GL2_3x_ErTv) + sd(gw_het_GL2_3x_ErTv), geom = "pointrange" , size=.12, stroke=.4, col="black") +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"), label.y = c(0.00024, 0.00020), size=2.5) +
  
  # embellish
  scale_y_continuous(limits=c(0.00007, 0.00025),labels = function(x) format(x, scientific = FALSE)) +
  
  
  labs(x=NULL, y="Genome-wide heterozygosity (tv)") + geom_text_repel(data=data, aes(label=lab_ID), segment.size=.15, size=2)

gw_het
```



### With GL2 <5x estimate excluding the 4 samples of average depth of coverage <5x

```{r}
# plot het values per time window
gw_het = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(time_window, gw_het_GL2_5x_ErTv)) + 
  theme_bw() + theme(legend.position = "none", axis.text=element_text(size=7, colour='gray10'), axis.title.y=element_text(size=7), 
                     panel.grid.major=(element_line(size=.12, colour='gray80')))
  
  
# plot violin per group
gw_het = gw_het +  geom_violin(aes(colour=time_window, fill=time_window), lwd=.3, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(colour='gray40', shape=16, position=position_jitter(0.11), cex=.6, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='black', width=.12, lwd=.3) +
  
  stat_summary(fun=mean, fun.min = function (gw_het_GL2_5x_ErTv) mean(gw_het_GL2_5x_ErTv) - sd(gw_het_GL2_5x_ErTv), 
               fun.max = function(gw_het_GL2_5x_ErTv) mean(gw_het_GL2_5x_ErTv) + sd(gw_het_GL2_5x_ErTv), geom = "pointrange" , size=.12, stroke=.4, col="black") +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"), label.y = c(0.00027, 0.00020), size=2.5) +
  
  # embellish
  scale_y_continuous(limits=c(0.00007, 0.00030),labels = function(x) format(x, scientific = FALSE)) +
  
  
  labs(x=NULL, y="Genome-wide heterozygosity (tv)") + geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=lab_ID), segment.size=.15, size=2)

gw_het
```


### Het estimates in regions annotated as exons

```{r}
# plot het values per time window
gw_het = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(time_window, exon_het_GL2_5x_ErTv)) + 
  theme_bw() + theme(legend.position = "none", axis.text=element_text(size=7, colour='gray10'), axis.title.y=element_text(size=7), 
                     panel.grid.major=(element_line(size=.12, colour='gray80')))
  
  
# plot violin per group
gw_het = gw_het +  geom_violin(aes(colour=time_window, fill=time_window), lwd=.3, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(colour='gray40', shape=16, position=position_jitter(0.11), cex=.6, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='black', width=.12, lwd=.3) +
  
  stat_summary(fun=mean, fun.min = function (exon_het_GL2_5x_ErTv) mean(exon_het_GL2_5x_ErTv) - sd(exon_het_GL2_5x_ErTv), 
               fun.max = function(exon_het_GL2_5x_ErTv) mean(exon_het_GL2_5x_ErTv) + sd(exon_het_GL2_5x_ErTv), geom = "pointrange" , size=.12, stroke=.4, col="black") +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"), label.y = c(0.0002, 0.00018), size=2.5) +
  
  # embellish
  #scale_y_continuous(limits=c(0.00007, 0.00030),labels = function(x) format(x, scientific = FALSE)) +
  
  
  labs(x=NULL, y="Heterozygosity at exons (tv)") + geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=lab_ID), segment.size=.15, size=2)

gw_het
```



### Results of ngsF (inbreeding coefficients per group)

```{r}
data$time_window=factor(data$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```

```{r}
# plot het values per time window
indFplot = ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TGF02' & lab_ID!='mwr_TGF09'), 
              aes(time_window, F_ngsF)) + 
  theme_bw() + theme(legend.position = "none", axis.text=element_text(size=7, colour='gray10'), axis.title.y=element_text(size=7), 
                     panel.grid.major=(element_line(size=.12, colour='gray80')))
  
  
# plot violin per group
indFplot = indFplot +  geom_violin(aes(colour=time_window, fill=time_window), lwd=.3, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(colour='gray40', shape=16, position=position_jitter(0.11), cex=.6, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='black', width=.12, lwd=.3) +
  
  stat_summary(fun=mean, fun.min = function (F_ngsF) mean(F_ngsF) - sd(F_ngsF), 
               fun.max = function(F_ngsF) mean(F_ngsF) + sd(F_ngsF), geom = "pointrange" , size=.12, stroke=.4, col="black") +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "greater"), label.y = c(0.00027, 0.00020), size=2.5) #+
  
  # embellish
  #scale_y_continuous(limits=c(0.00007, 0.00030),labels = function(x) format(x, scientific = FALSE)) +
  
  
#  labs(x=NULL, y="Inbreeding coefficient") + geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TGF02' & lab_ID!='mwr_TGF09'), aes(label=lab_ID), segment.size=.15, size=2)

indFplot
```




#############################################################################

## FRoH

```{r}
data$time_window=factor(data$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```

```{r}
# plot het values per time window

#ggplot(data, aes(time_window, Froh_10scf)) +

#ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'),
 #    aes(time_window, Froh_10scf)) +
  
#ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TG0452' & lab_ID!='mwr_TG0461'),
 #      aes(time_window, Froh_10scf)) +

ggplot(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01' & lab_ID!='mwr_TGF02' & lab_ID!='mwr_TGF04' & lab_ID!='mwr_TGF09' & lab_ID!='mwr_TG0452' & lab_ID!='mwr_TG0461'),
       aes(time_window, Froh_10scf)) +


# set up the panel
theme_bw() + theme(legend.position = "none", axis.text=element_text(size=7, colour='gray10'), axis.title.y=element_text(size=7), 
                     panel.grid.major=(element_line(size=.12, colour='gray80'))) +

# plot violin per group
geom_violin(aes(colour=time_window, fill=time_window), lwd=.3, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(colour='gray40', shape=16, position=position_jitter(0.11), cex=.6, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='black', width=.12, lwd=.3) +
  
  stat_summary(fun=mean, fun.min = function (Froh_10scf) mean(Froh_10scf) - sd(Froh_10scf), 
               fun.max = function(Froh_10scf) mean(Froh_10scf) + sd(Froh_10scf), geom = "pointrange" , size=.12, stroke=.4, col="black") +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "less")) +
  
  geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=lab_ID), segment.size=.15, size=2)
  

```



### With filtered vcf files



```{r}
data$time_window=factor(data$time_window,levels=c("NWRpre","NWRpost","SWRpre", "SWRpost"))
```

```{r}
# plot het values per time window

ggplot(data, aes(time_window, Froh_88filt)) +

# set up the panel
theme_bw() + theme(legend.position = "none", axis.text=element_text(size=7, colour='gray10'), axis.title.y=element_text(size=7), 
                     panel.grid.major=(element_line(size=.12, colour='gray80'))) +

# plot violin per group
geom_violin(aes(colour=time_window, fill=time_window), lwd=.3, alpha=.4) +
  
  # add scatter points per group
  geom_jitter(colour='gray40', shape=16, position=position_jitter(0.11), cex=.6, alpha=.7) + 
  
  # add colours and filling colours
  scale_fill_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) + 
  scale_color_manual(values= (c("gold3", "darkorange", "royalblue4", "steelblue3"))) +
  
  # add stats and an unpaired wilcox test to compare the medians
  stat_summary(fun=median, geom="crossbar", colour='black', width=.12, lwd=.3) +
  
  stat_summary(fun=mean, fun.min = function (Froh_10scf) mean(Froh_10scf) - sd(Froh_10scf), 
               fun.max = function(Froh_10scf) mean(Froh_10scf) + sd(Froh_10scf), geom = "pointrange" , size=.12, stroke=.4, col="black") +
  
  stat_compare_means(paired=F, method = "wilcox.test", comparisons = list(c("NWRpre", "NWRpost"), c("SWRpre", "SWRpost")), method.args = list(alternative = "less")) +
  
  geom_text_repel(data=subset(data, lab_ID!='mwr_TG0450' & lab_ID!='mwr_TG0451' & lab_ID!='mwr_TG0456' & lab_ID!='mwr_TGF01'), aes(label=lab_ID), segment.size=.15, size=2)
  

```























