---
title: "WR_v4_relatedness"
author: "Fátima S. Barreiro"
date: "November 23, 2020"
output: html_document
---




## Computing pairwise relatedness with ngsRelate


```{bash}
########################################
# Relatedness test for each subspecies 
######################


## Input needed: genotype likelihoods as log scaled GL in binary format (.glf), frequencies (.freq) and sample names (IDlist.txt)


# Make .glf file from .beagle file

zcat NWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.beagle.gz | tail -n +2 | perl -an -e 'for($i=3; $i<=$#F; $i++){print(pack("d",($F[$i]==0 ? -inf : log($F[$i]))))}'  > NWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.glf

zcat SWR.27I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.beagle.gz | tail -n +2 | perl -an -e 'for($i=3; $i<=$#F; $i++){print(pack("d",($F[$i]==0 ? -inf : log($F[$i]))))}'  > SWR.27I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.glf

## before using it I checked that its size in bytes is equal to (N_IND * N_SITES * 8 * 3)

## compress the glf files
gzip NWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.glf
gzip SWR.27I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.glf


# Make .freq file from .maf file
zcat NWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.mafs.gz | cut -f6 | sed 1d > NWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.freq

zcat SWR.27I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.mafs.gz | cut -f6 | sed 1d > SWR.27I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.freq


# Make ID file
cut -d '/' -f9 NWR_total_bamlist.txt | cut -d . -f1 > NWR_IDlist.txt
cut -d '/' -f9 SWR_total_bamlist.txt | cut -d . -f1 > SWR_IDlist.txt


# Run ngsRelate
ngsRelate -g NWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.glf.gz  -f NWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.freq -n 25 -z NWR_IDlist.txt -O NWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.res

ngsRelate -g SWR.27I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.glf.gz  -f SWR.27I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.freq -n 27 -z SWR_IDlist.txt -O SWR.27.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.res
```





## Parsing output and visualizing results

Load required libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(cowplot)
library(ggrepel)
library(reshape2)
library(dplyr)
library(tidyverse)
library(data.table)
require(scales)
library(ggpubr)
library(tidyr)
library(PairedData)
library(gridExtra)
```



Open tables of pairwise relatedness (output of ngsRelate) per subspecies
```{r}
nwr = read.table("NWR.25I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.res", header=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
swr = read.table("SWR.27I.above13Mb_NoSexChr_NonMaskedRegions_GL2.tv.res", header=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```



### Visualize the pairwise relatedness coefficients for each subspecies

#### NWR
```{r}
ggplot(nwr, aes(R1, R0)) + geom_point (alpha=.8, color="darkorange", size=2) + theme_bw() + labs(x="R1", y="R0", title="NWR pairwise relatedness among samples") 
  #scale_x_continuous(limits=c(0,1))

ggplot(nwr, aes(R1, KING)) + geom_point (alpha=.8, color="darkorange", size=2) + theme_bw() + labs(x="R1", y="KING", title="NWR pairwise relatedness among samples") 
  #scale_x_continuous(limits=c(0,1))
```
One related pair in NWR.



#### SWR
```{r}
ggplot(swr, aes(R1, R0)) + geom_point (alpha=.8, color="royalblue4", size=2) + theme_bw() + labs(x="R1", y="R0", title="SWR pairwise relatedness among samples") 
  #scale_x_continuous(limits=c(0,1))

ggplot(swr, aes(R1, KING)) + geom_point (alpha=.8, color="royalblue4", size=2) + theme_bw() + labs(x="R1", y="KING", title="SWR pairwise relatedness among samples") 
  #scale_x_continuous(limits=c(0,1))
```
Two related pairs in SWR.



### Find out who the relatives are in each subspecies (outlier pairs in the plots above)
```{r}
nwr_rels = subset(nwr, R1 > 0.6)

swr_rels = subset(swr, R1 > 10)

# these lines represent the pairs of samples with signal of relatedness
head(nwr_rels)
head(swr_rels)
```




### Find the sample names (not lab IDs) corresponding to those related pairs

Open metadata and genomic metrics table
```{r}
md = read.csv("WR_TableS1_v6.csv", h=T, fill=T, stringsAsFactors=T, fileEncoding="latin1")
```


```{r}
md[md$lab_ID %in% as.vector(nwr_rels$ida),]
md[md$lab_ID %in% as.vector(nwr_rels$idb),]

md[md$lab_ID %in% as.vector(swr_rels$ida),]
md[md$lab_ID %in% as.vector(swr_rels$idb),]


```


Change old 'sample_name' by the new_IDs for plotting
```{r}
# NWR
nwr$ida <- replace(as.character(nwr$ida), nwr$ida == "mwr_TG0450", "CD-un.1")
nwr$idb <- replace(as.character(nwr$idb), nwr$idb == "mwr_TG0458", "CD-un.2")


# SWR
## pair 1
swr$ida <- replace(as.character(swr$ida), swr$ida == "mwr_TGF02", "un1856.1")
swr$idb <- replace(as.character(swr$idb), swr$idb == "mwr_TGF03", "ZA-un.1")

swr$ida <- replace(as.character(swr$ida), swr$ida == "mwr_TGF04", "ZA1845.2")
swr$idb <- replace(as.character(swr$idb), swr$idb == "mwr_TGF09", "ZA1842.1")

```


```{r}
nwr_rels = subset(nwr, R1 > 0.6)

swr_rels = subset(swr, R1 > 10)

# these lines represent the pairs of samples with signal of relatedness
head(nwr_rels)
head(swr_rels)
```


### Plot again now with the labels on the related individuals

#### NWR
```{r}
nwr0 = ggplot(nwr, aes(R1, R0)) + geom_point (alpha=.8, color="darkorange", size=2) + theme_bw() + labs(x="R1", y="R0") +
  geom_text_repel(data=subset(nwr, R1 > 0.6), aes(label=ida), segment.size=.15, size=2, colour="gray20", fontface="bold", point.padding=.5) +
  geom_text_repel(data=subset(nwr, R1 > 0.6), aes(label=idb), segment.size=.15, size=2, colour="gray20", fontface="bold", point.padding=.5, xlim=c(NULL, 0.8), ylim=c(0, 0.2)) 

nwrK = ggplot(nwr, aes(R1, KING)) + geom_point (alpha=.8, color="darkorange", size=2) + theme_bw() + labs(x="R1", y="KING") +
    geom_text_repel(data=subset(nwr, R1 > 0.6), aes(label=ida), segment.size=.15, size=2, colour="gray20", fontface="bold", point.padding=.5) +
    geom_text_repel(data=subset(nwr, R1 > 0.6), aes(label=idb), segment.size=.15, size=2, colour="gray20", fontface="bold", point.padding=.5, xlim=c(NULL, 0.8), ylim=c(0.2, 0.3)) 

nwr0
nwrK

```




#### SWR
```{r}
swr0 = ggplot(swr, aes(R1, R0)) + geom_point (alpha=.8, color="royalblue4", size=2) + theme_bw() + labs(x="R1", y="R0") +
  scale_x_continuous(breaks=seq(0,16,2)) +
   geom_text_repel(data=subset(swr, R1 > 10), aes(label=ida), segment.size=.15, size=2, colour="gray20", fontface="bold", point.padding=.5, direction="x", xlim=c(8,14), ylim=c(0, 0.5)) +
  geom_text_repel(data=subset(swr, R1 > 10), aes(label=idb), segment.size=.15, size=2, colour="gray20", fontface="bold", point.padding=.5, direction="y", ylim=c(0.15,0.5))

swrK = ggplot(swr, aes(R1, KING)) + geom_point (alpha=.8, color="royalblue4", size=2) + theme_bw() + labs(x="R1", y="KING") +
  scale_x_continuous(breaks=seq(0,16,2)) +
  geom_text_repel(data=subset(swr, R1 > 10), aes(label=ida), segment.size=.15, size=2, colour="gray20", fontface="bold", point.padding=.5, direction="x", xlim=c(7, 13), ylim=c(0.35,0.4)) +
  geom_text_repel(data=subset(swr, R1 > 10), aes(label=idb), segment.size=.15, size=2, colour="gray20", fontface="bold", point.padding=.5, direction="both", xlim=c(9,14), ylim=c(0.1,0.3))

swr0
swrK
```


```{r}
plots = ggarrange(nwr0, nwrK, swr0, swrK, nrow=2, ncol=2, labels=c("A", "B" ,"C", "D"), align="v")
plots
```



```{r}
pdf("WR_v4_FigS5_Relatedness", useDingbats = F)
plots
dev.off()
```













